---
title: "Critical Thinking - Behavioral Analysis"
author: "Jiayin Zhi"
date: "2025-08-25"
output: 
  html_document:
    code_folding: hide    
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged   
---

# Data
```{r}
library(tidyverse)
library(readxl)
library(janitor)

script_dir <- dirname(rstudioapi::getSourceEditorContext()$path)
read_first_sheet <- function(path) {
  read_excel(path, sheet = 1) %>% clean_names()
}

df_main <- read_first_sheet(file.path(script_dir, "study_data", "main_study_data.xlsx"))
df_argumentation <- read_first_sheet(file.path(script_dir, "behavioral_log_based_data", "argument_and_referenced_docs_overlap.xlsx"))
df_copy_overlap <- read_first_sheet(file.path(script_dir, "behavioral_log_based_data", "copy_rate_and_textual_overlap.xlsx"))
df_task_approach <- read_first_sheet(file.path(script_dir, "behavioral_log_based_data", "task_approaches.xlsx"))
df_early <- read_first_sheet(file.path(script_dir, "behavioral_log_based_data", "early_access_data.xlsx"))
df_late <- read_first_sheet(file.path(script_dir, "behavioral_log_based_data", "late_access_data.xlsx"))
```

## Helper Functions
```{r}
derive_factors <- function(data) {
  data %>%
    mutate(
      condition = factor(condition,
        levels = c("10_control","10_continuous","10_early","10_late",
                   "30_control","30_continuous","30_early","30_late")),
      time_availability = factor(
        case_when(str_starts(as.character(condition), "10") ~ "Insufficient",
                  str_starts(as.character(condition), "30") ~ "Sufficient"),
        levels = c("Insufficient","Sufficient")),
      ai_access = factor(
        case_when(str_detect(as.character(condition), "early")      ~ "EarlyAI",
                  str_detect(as.character(condition), "continuous") ~ "ContinuousAI",
                  str_detect(as.character(condition), "late")       ~ "LateAI",
                  str_detect(as.character(condition), "control")    ~ "NoAI"),
        levels = c("EarlyAI","ContinuousAI","LateAI","NoAI"))
    )
}

print_by_category <- function(desc, vars, label) {
  cat("\n---", label, "---\n")
  for (var in vars) {
    cat("\n", var, ":\n")
    desc %>% filter(measure == var) %>% select(-measure) %>% print()
  }
}
```

# Argument and Document Reference Overlap
This section analyzes the overlap between participants' essays and AI responses in terms of:
- **Arguments**: The number of arguments in participants' essays that overlap with AI-generated arguments vs. non-overlapping arguments
- **Document references**: The documents referenced in participants' essays vs. documents referenced by the AI (overlapping/non-overlapping, viewed/unviewed)
```{r}
argumentation_vars <- c(
  # Document reference metrics
  "viewed_doc_referenced_count",    # docs referenced in essay that participant viewed
  "viewed_doc_referenced_overlap",  # overlap between viewed docs and AI-refered docs
  "num_unviewed_doc_referenced",    # docs referenced in essay that participant did not view
  "overlap_doc_reference_count",    # docs referenced in both essay and AI responses
  "overlap_doc_reference",          # proportion of overlapping doc references
  "num_new_doc_referenced",         # docs referenced in essay but not in AI responses
  # Argument metrics
  "essay_argument_count",           # total arguments in participant's essay
  "ai_argument_count",              # total arguments in AI responses
  "overlap_arguments_count",        # arguments appearing in both essay and AI responses
  "num_new_arguments"               # arguments in essay not found in AI responses
)

df_argumentation_analysis <- df_argumentation %>%
  mutate(across(any_of(argumentation_vars), ~ replace_na(., 0))) %>%
  derive_factors()

# Descriptive statistics (overall)
argumentation_desc_overall <- df_argumentation_analysis %>%
  select(all_of(argumentation_vars)) %>%
  pivot_longer(everything(), names_to = "measure", values_to = "value") %>%
  group_by(measure) %>%
  summarise(N = n(), Mean = mean(value), SD = sd(value),
            Median = median(value), Min = min(value), Max = max(value), .groups = "drop")

# Descriptive statistics (by factorial cell)
argumentation_desc_by_factors <- df_argumentation_analysis %>%
  select(time_availability, ai_access, all_of(argumentation_vars)) %>%
  pivot_longer(-c(time_availability, ai_access), names_to = "measure", values_to = "value") %>%
  group_by(time_availability, ai_access, measure) %>%
  summarise(N = n(), Mean = mean(value), SD = sd(value),
            Median = median(value), Min = min(value), Max = max(value), .groups = "drop")

# Summaries by categories of metrics
doc_ref_in_essay_vars <- c("viewed_doc_referenced_count", "viewed_doc_referenced_overlap", "num_unviewed_doc_referenced")
doc_ref_overlap_vars <- c("overlap_doc_reference_count", "overlap_doc_reference", "num_new_doc_referenced")
argument_overlap_vars <- c("essay_argument_count", "ai_argument_count", "overlap_arguments_count", "num_new_arguments")

print_by_category <- function(desc, vars, label) {
  cat("\n---", label, "---\n")
  for (var in vars) {
    cat("\n", var, ":\n")
    desc %>% filter(measure == var) %>% select(-measure) %>% print()
  }
}

# Print the results
print(argumentation_desc_overall)
print_by_category(argumentation_desc_by_factors, doc_ref_in_essay_vars, "Documents Referenced in Essays")
print_by_category(argumentation_desc_by_factors, doc_ref_overlap_vars, "Document Reference Overlap (Participants' Essays vs. AI Responses)")
print_by_category(argumentation_desc_by_factors, argument_overlap_vars, "Argument Overlap (Participants' Essays vs. AI Responses)")
```

# Document Engagement, AI Interaction, and Writing Behavior
This section reports behavioral metrics from the main task interface, including:
- **AI interaction**: Number of queries sent to the LLM and proportion of time spent in the AI panel
- **Writing process**: Essay word count, writing pauses, and proportion of time spent writing
- **Document engagement**: Documents viewed/clicked before and after writing began
- **Time allocation**: Proportion of task time spent in each interface panel (document viewer, AI chatbot, essay editor, other)
```{r}
main_behavioral_vars <- c(
  # AI interaction
  "ai_total_queries",               # total queries sent to the LLM
  "perc_ai_time",                   # % of task time in the AI panel
  # Writing process
  "essay_word_count",               # final essay word count
  "sum_writing_pause_times",        # total pause duration during writing (ms)
  "rate_pause_during_writing",      # pause rate during writing
  "perc_essay_time",                # % of task time in the essay panel
  # Document engagement
  "num_doc_viewed_before_writing",  # unique docs viewed before writing began
  "num_doc_clicked_before_writing", # doc clicks before writing began
  "prec_time_before_writing",       # % of task time spent before writing began
  "num_doc_viewed_after_writing",   # unique docs viewed after writing began
  "num_doc_further_viewed",         # docs further viewed during writing
  "num_doc_further_clicked",        # doc clicks during writing
  "rate_doc_further_viewed",        # rate of further doc viewing during writing
  # Time allocation
  "perc_doc_time",                  # % of task time in the document viewer
  "perc_essay_time",                # % of task time in the essay panel
  "perc_ai_time",                   # % of task time in the AI panel
  "perc_rest_time"                  # % of task time in other areas
)

df_main_behavioral <- df_main %>%
  mutate(across(any_of(main_behavioral_vars), ~ replace_na(as.numeric(.), 0))) %>%
  derive_factors()

# Descriptive statistics (overall)
main_behavioral_desc_overall <- df_main_behavioral %>%
  select(all_of(main_behavioral_vars)) %>%
  pivot_longer(everything(), names_to = "measure", values_to = "value") %>%
  group_by(measure) %>%
  summarise(N = n(), Mean = mean(value), SD = sd(value),
            Median = median(value), Min = min(value), Max = max(value), .groups = "drop")

# Descriptive statistics (by factorial cell)
main_behavioral_desc_by_factors <- df_main_behavioral %>%
  select(time_availability, ai_access, all_of(main_behavioral_vars)) %>%
  pivot_longer(-c(time_availability, ai_access), names_to = "measure", values_to = "value") %>%
  group_by(time_availability, ai_access, measure) %>%
  summarise(N = n(), Mean = mean(value), SD = sd(value),
            Median = median(value), Min = min(value), Max = max(value), .groups = "drop")

# Summaries by categories of metrics
ai_interaction_vars <- c("ai_total_queries", "perc_ai_time")
writing_process_vars <- c("essay_word_count", "sum_writing_pause_times", "rate_pause_during_writing", "perc_essay_time")
doc_engagement_vars <- c("num_doc_viewed_before_writing", "num_doc_clicked_before_writing",
                         "prec_time_before_writing", "num_doc_viewed_after_writing",
                         "rate_doc_further_viewed", "num_doc_further_clicked", "num_doc_further_viewed")
time_allocation_vars <- c("perc_doc_time", "perc_ai_time", "perc_essay_time", "perc_rest_time")

# Print the results
print(main_behavioral_desc_overall)
print_by_category(main_behavioral_desc_by_factors, ai_interaction_vars, "AI Interaction")
print_by_category(main_behavioral_desc_by_factors, writing_process_vars, "Writing Process")
print_by_category(main_behavioral_desc_by_factors, doc_engagement_vars, "Document Engagement")
print_by_category(main_behavioral_desc_by_factors, time_allocation_vars, "Time Allocation")
```

# Early and Late LLM Access Specific Analysis
This section reports behavioral metrics specific to the Early and Late LLM access conditions, focusing on document engagement patterns relative to when the LLM was introduced:
- **Early LLM access**: Document engagement during and after the LLM was introduced
- **Late LLM access**: Document engagement before and during the LLM was introduced, plus essay word count changes after LLM introduction
```{r}
# Early AI access (10_early and 30_early)
early_vars <- c(
  "num_doc_clicked_during_injection",   # doc clicks while LLM was being introduced
  "num_doc_viewed_during_injection",    # unique docs viewed while LLM was being introduced
  "num_doc_clicked_after_injection",    # doc clicks after LLM introduction
  "num_doc_viewed_after_injection",     # unique docs viewed after LLM introduction
  "rate_doc_after_injection_viewed",    # rate of doc viewing after LLM introduction
  "rate_doc_after_injection_clicked",   # rate of doc clicking after LLM introduction
  "num_doc_early_further_clicked",      # further doc clicks in early access condition
  "num_doc_early_further_viewed"        # further docs viewed in early access condition
)

df_early_analysis <- df_early %>%
  mutate(
    across(any_of(early_vars), ~ replace_na(., 0)),
    condition = factor(condition, levels = c("10_early", "30_early")),
    time_availability = factor(
      case_when(str_starts(as.character(condition), "10") ~ "Insufficient",
                str_starts(as.character(condition), "30") ~ "Sufficient"),
      levels = c("Insufficient", "Sufficient"))
  )

early_desc_by_condition <- df_early_analysis %>%
  select(condition, time_availability, all_of(early_vars)) %>%
  pivot_longer(-c(condition, time_availability), names_to = "measure", values_to = "value") %>%
  group_by(condition, time_availability, measure) %>%
  summarise(N = n(), Mean = mean(value), SD = sd(value),
            Median = median(value), Min = min(value), Max = max(value), .groups = "drop")

# Late AI access (10_late and 30_late)
late_vars <- c(
  "num_doc_clicked_before_late_injection",   # doc clicks before LLM introduction
  "num_doc_viewed_before_late_injection",    # unique docs viewed before LLM introduction
  "num_doc_clicked_during_late_injection",   # doc clicks while LLM was being introduced
  "num_doc_viewed_during_late_injection",    # unique docs viewed while LLM was being introduced
  "rate_doc_during_late_injection_viewed",   # rate of doc viewing during LLM introduction
  "rate_doc_during_late_injection_clicked",  # rate of doc clicking during LLM introduction
  "num_doc_late_further_clicked",            # further doc clicks in late access condition
  "num_doc_late_further_viewed"              # further docs viewed in late access condition
)

df_late_analysis <- df_late %>%
  mutate(
    across(any_of(late_vars), ~ replace_na(., 0)),
    timeout_essay_wordcount = ifelse(is.na(timeout_essay_text) | timeout_essay_text == "", 0,
                                     sapply(strsplit(as.character(timeout_essay_text), "\\s+"), length)),
    final_essay_wordcount = ifelse(is.na(essay_text) | essay_text == "", 0,
                                   sapply(strsplit(as.character(essay_text), "\\s+"), length)),
    wordcount_difference = final_essay_wordcount - timeout_essay_wordcount,
    condition = factor(condition, levels = c("10_late", "30_late")),
    time_availability = factor(
      case_when(str_starts(as.character(condition), "10") ~ "Insufficient",
                str_starts(as.character(condition), "30") ~ "Sufficient"),
      levels = c("Insufficient", "Sufficient"))
  )

late_vars_with_wordcount <- c(late_vars,
  "timeout_essay_wordcount",    # essay word count at the time LLM was introduced
  "final_essay_wordcount",      # final essay word count
  "wordcount_difference"        # change in word count after LLM introduction
)

late_desc_by_condition <- df_late_analysis %>%
  select(condition, time_availability, all_of(late_vars_with_wordcount)) %>%
  pivot_longer(-c(condition, time_availability), names_to = "measure", values_to = "value") %>%
  group_by(condition, time_availability, measure) %>%
  summarise(N = n(), Mean = mean(value), SD = sd(value),
            Median = median(value), Min = min(value), Max = max(value), .groups = "drop")

# Print the results
cat("\n--- Early AI Access ---\n")
for (var in early_vars) {
  cat("\n", var, ":\n")
  early_desc_by_condition %>% filter(measure == var) %>% select(-measure) %>% print()
}

cat("\n--- Late AI Access ---\n")
for (var in late_vars_with_wordcount) {
  cat("\n", var, ":\n")
  late_desc_by_condition %>% filter(measure == var) %>% select(-measure) %>% print()
}
```

# Plots and Tables
## Plot of Argument Overlap
```{r}
library(ggpattern)

arg_data <- data.frame(
  Time = rep(c("Insufficient", "Sufficient"), each = 4),
  AI_Access = rep(c("Early LLM access", "Continuous LLM access", "Late LLM access", "No LLM access"), 2),
  N = c(39, 50, 50, 51, 51, 52, 48, 52),
  AI_Arguments  = c(1.15, 1.46, 0.78, 0.00, 1.71, 2.48, 1.58, 0.00),
  New_Arguments = c(2.56, 1.80, 1.80, 2.98, 2.84, 1.87, 3.60, 5.08),
  SD_AI  = c(1.51, 1.73, 1.49, 0.00, 1.65, 2.19, 1.90, 0.00),
  SD_New = c(2.17, 1.74, 1.70, 2.04, 1.99, 1.93, 2.48, 1.94)
) %>%
  mutate(SE_AI = SD_AI / sqrt(N), SE_New = SD_New / sqrt(N),
         AI_Access = factor(AI_Access, levels = c("Early LLM access", "Continuous LLM access", "Late LLM access", "No LLM access")))

arg_long <- arg_data %>%
  pivot_longer(c(AI_Arguments, New_Arguments), names_to = "Source", values_to = "Count") %>%
  mutate(Source = factor(Source, levels = c("AI_Arguments", "New_Arguments"),
                         labels = c("Arguments overlap with the LLM", "Unique arguments")))

se_top <- bind_rows(
  arg_data %>% transmute(Time, AI_Access,
    Source = factor("Unique arguments", levels = c("Arguments overlap with the LLM", "Unique arguments")),
    y_top = New_Arguments, SE = SE_New),
  arg_data %>% transmute(Time, AI_Access,
    Source = factor("Arguments overlap with the LLM", levels = c("Arguments overlap with the LLM", "Unique arguments")),
    y_top = New_Arguments + AI_Arguments, SE = SE_AI)
) %>% mutate(ymin = y_top - SE, ymax = y_top + SE)

fill_cols <- c("Arguments overlap with the LLM" = "#F4A3A8", "Unique arguments" = "#A8DADC")
patterns <- c("Arguments overlap with the LLM" = "stripe", "Unique arguments" = "none")

p_args <- ggplot(arg_long, aes(x = Time, y = Count, fill = Source, pattern = Source)) +
  geom_bar_pattern(stat = "identity", width = 0.5, color = "black", linewidth = 0.4,
                   pattern_fill = "white", pattern_colour = "white", pattern_angle = 45,
                   pattern_density = 0.1, pattern_spacing = 0.03, pattern_key_scale_factor = 1) +
  geom_errorbar(data = se_top, aes(x = Time, y = y_top, ymin = ymin, ymax = ymax),
                inherit.aes = FALSE, width = 0.16, linewidth = 0.4, alpha = 0.9) +
  geom_text(aes(label = ifelse(Count > 0.15, round(Count, 1), "")),
            position = position_stack(vjust = 0.5), size = 3.5, color = "black", fontface = "bold") +
  facet_grid(~ AI_Access) +
  scale_fill_manual(values = fill_cols) +
  scale_pattern_manual(values = patterns, guide = "none") +
  guides(fill = guide_legend(override.aes = list(pattern = c("stripe", "none")))) +
  coord_cartesian(ylim = c(0, 6)) +
  theme_bw() +
  theme(legend.position = "top", panel.spacing = unit(1.5, "lines"),
        panel.grid.minor = element_blank(), panel.grid.major.x = element_blank(),
        strip.text = element_text(size = 11, face = "bold"),
        axis.text.x = element_text(size = 10, face = "bold")) +
  labs(x = "Time availability", y = "Number of arguments", fill = NULL)

print(p_args)

ggsave(file.path("behavioral_engagement_plots", "argument_composition_by_ai_access.pdf"),
       p_args, width = 12, height = 4.5, device = "pdf")
```

## Tables of Textual Overlap
```{r}
# Copying: participants copying text from AI responses into their essays
copy_table <- df_copy_overlap %>%
  group_by(condition) %>%
  summarise(
    N = n(),
    n_copied = sum(copied_wordcount > 0, na.rm = TRUE),
    `Instances (%)` = paste0(n_copied, " (", round(100 * n_copied / n(), 0), "%)"),
    Mean_rate_all = round(mean(copied_rate, na.rm = TRUE) * 100, 2),
    SD_rate_all = round(sd(copied_rate, na.rm = TRUE) * 100, 2),
    `Avg Rate - All (M±SD)` = paste0(Mean_rate_all, " ± ", SD_rate_all),
    Mean_rate_copiers = round(mean(copied_rate[copied_wordcount > 0], na.rm = TRUE) * 100, 2),
    SD_rate_copiers = round(sd(copied_rate[copied_wordcount > 0], na.rm = TRUE) * 100, 2),
    `Avg Rate - Copiers (M±SD)` = paste0(
      ifelse(is.na(Mean_rate_copiers), "–", Mean_rate_copiers), " ± ",
      ifelse(is.na(SD_rate_copiers), "–", SD_rate_copiers)),
    Mean_copied = round(mean(copied_wordcount[copied_wordcount > 0], na.rm = TRUE), 1),
    SD_copied = round(sd(copied_wordcount[copied_wordcount > 0], na.rm = TRUE), 1),
    `Wordcount (M±SD)` = paste0(
      ifelse(is.na(Mean_copied), "–", Mean_copied), " ± ",
      ifelse(is.na(SD_copied), "–", SD_copied)),
    .groups = "drop"
  ) %>%
  select(condition, N, `Instances (%)`, `Avg Rate - All (M±SD)`, `Avg Rate - Copiers (M±SD)`, `Wordcount (M±SD)`) %>%
  arrange(condition)

# Textual overlap: overlap between AI responses and participant essays (not necessarily copied)
overlap_table <- df_copy_overlap %>%
  group_by(condition) %>%
  summarise(
    N = n(),
    n_overlap = sum(overlap_wordcount > 0, na.rm = TRUE),
    `Instances (%)` = paste0(n_overlap, " (", round(100 * n_overlap / n(), 0), "%)"),
    Mean_rate_all = round(mean(overlap_rate, na.rm = TRUE) * 100, 2),
    SD_rate_all = round(sd(overlap_rate, na.rm = TRUE) * 100, 2),
    `Avg Rate - All (M±SD)` = paste0(Mean_rate_all, " ± ", SD_rate_all),
    Mean_rate_overlap = round(mean(overlap_rate[overlap_wordcount > 0], na.rm = TRUE) * 100, 2),
    SD_rate_overlap = round(sd(overlap_rate[overlap_wordcount > 0], na.rm = TRUE) * 100, 2),
    `Avg Rate - Overlap Cases (M±SD)` = paste0(
      ifelse(is.na(Mean_rate_overlap), "–", Mean_rate_overlap), " ± ",
      ifelse(is.na(SD_rate_overlap), "–", SD_rate_overlap)),
    Mean_overlap = round(mean(overlap_wordcount[overlap_wordcount > 0], na.rm = TRUE), 1),
    SD_overlap = round(sd(overlap_wordcount[overlap_wordcount > 0], na.rm = TRUE), 1),
    `Wordcount (M±SD)` = paste0(
      ifelse(is.na(Mean_overlap), "–", Mean_overlap), " ± ",
      ifelse(is.na(SD_overlap), "–", SD_overlap)),
    .groups = "drop"
  ) %>%
  select(condition, N, `Instances (%)`, `Avg Rate - All (M±SD)`, `Avg Rate - Overlap Cases (M±SD)`, `Wordcount (M±SD)`) %>%
  arrange(condition)

cat("\n--- Copying ---\n")
print(copy_table)

cat("\n--- Textual Overlap ---\n")
print(overlap_table)
```

## Plot of Iterative Document Engagement
```{r}
library(ggpattern)

doc_data <- data.frame(
  Time = rep(c("Insufficient", "Sufficient"), each = 4),
  AI_Access = rep(c("Early LLM access", "Continuous LLM access", "Late LLM access", "No LLM access"), 2),
  N = c(39, 50, 50, 51, 51, 52, 48, 52),
  Docs_Before = c(3.54, 3.46, 3.42, 4.04, 4.41, 3.75, 4.77, 4.96),
  SD_Before = c(2.49, 2.54, 2.72, 2.77, 2.48, 2.96, 2.82, 2.93),
  Docs_During = c(1.51, 0.84, 1.82, 1.55, 2.70, 2.22, 3.58, 3.63),
  SD_During = c(1.99, 1.43, 2.14, 1.92, 2.96, 2.55, 2.59, 2.92)
) %>%
  mutate(SE_Before = SD_Before / sqrt(N), SE_During = SD_During / sqrt(N),
         AI_Access = factor(AI_Access, levels = c("Early LLM access", "Continuous LLM access", "Late LLM access", "No LLM access")))

doc_data_long <- doc_data %>%
  pivot_longer(c(Docs_Before, Docs_During), names_to = "Timing", names_pattern = "Docs_(.*)", values_to = "Count") %>%
  mutate(SE = ifelse(Timing == "Before", SE_Before, SE_During),
         Timing = factor(Timing, levels = c("Before", "During"), labels = c("Before writing", "During writing")),
         Pattern = ifelse(Timing == "During writing", "stripe", "none"))

p_doc_viewing <- ggplot(doc_data_long, aes(x = AI_Access, y = Count, fill = Timing, pattern = Pattern)) +
  geom_bar_pattern(stat = "identity", position = position_dodge(width = 0.8), width = 0.68,
                   color = "black", linewidth = 0.3, pattern_fill = "white", pattern_colour = "white",
                   pattern_density = 0.3, pattern_spacing = 0.02) +
  scale_pattern_manual(values = c("none" = "none", "stripe" = "stripe"), guide = "none") +
  geom_errorbar(aes(ymin = Count - SE, ymax = Count + SE),
                position = position_dodge(width = 0.8), width = 0.25, alpha = 0.7) +
  facet_grid(~ Time) +
  scale_fill_manual(values = c("Before writing" = "#4DB6AC", "During writing" = "#E57373")) +
  guides(fill = guide_legend(override.aes = list(pattern = c("none", "stripe")))) +
  theme_bw() +
  theme(legend.position = "top", panel.spacing = unit(2, "lines"),
        panel.grid.minor = element_blank(), panel.grid.major.x = element_blank(),
        text = element_text(size = 11), strip.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5),
        plot.margin = margin(10, 15, 15, 15)) +
  labs(x = "LLM access timing", y = "Number of documents viewed", fill = NULL) +
  ylim(0, 6)

print(p_doc_viewing)

ggsave(file.path("behavioral_engagement_plots", "doc_viewing_iteration.pdf"),
       p_doc_viewing, width = 12.5, height = 4.5, device = "pdf")
```

# Summaries of Task Approaches
```{r}
code_labels <- c(
  "1" = "Confidence & Reassurance", "2" = "Verification & Clarification",
  "3" = "Summarizing Documents", "4" = "Structuring Pros & Cons",
  "5" = "Content Adoption", "6" = "Polishing & Refinement",
  "7" = "Minimal/No AI Engagement", "8" = "Systematic Document Review",
  "9" = "Selective Document Review", "10" = "Time Management Strategy",
  "11" = "Heuristics/Judgments"
)

df_expanded <- df_task_approach %>%
  separate_rows(task_approach, sep = ";") %>%
  mutate(task_approach = as.numeric(task_approach)) %>%
  filter(!is.na(task_approach)) %>%
  mutate(code_label = code_labels[as.character(task_approach)])

# Full distribution by condition
code_distribution <- df_expanded %>%
  group_by(condition, code_label) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(condition) %>%
  mutate(percentage = round(count / sum(count) * 100, 1)) %>%
  arrange(condition, desc(percentage))

cat("\n--- Distribution by Condition ---\n")
for (cond in unique(code_distribution$condition)) {
  cat("\n", cond, ":\n")
  code_distribution %>%
    filter(condition == cond) %>%
    select(code_label, count, percentage) %>%
    print(n = Inf)
}

# Top 3 task approaches per factorial cell
cat("\n--- Top 3 Task Approaches by Time × AI Access ---\n")
df_expanded %>%
  mutate(
    time = ifelse(grepl("^10_", condition), "Insufficient", "Sufficient"),
    ai_access = case_when(
      grepl("_early$", condition) ~ "Early LLM",
      grepl("_continuous$", condition) ~ "Continuous LLM",
      grepl("_late$", condition) ~ "Late LLM",
      grepl("_control$", condition) ~ "No LLM"
    )
  ) %>%
  group_by(time, ai_access, code_label) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(time, ai_access) %>%
  mutate(percentage = round(count / sum(count) * 100, 1)) %>%
  slice_max(percentage, n = 3) %>%
  arrange(time, ai_access, desc(percentage)) %>%
  print(n = Inf)
```




